---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# r4sda

- The goal of r4sda is to make it easier to work with large scale assessment studies data.
- Is a collection of simple, wrapper functions, mainly, generated to resolve common tasks when working with large scale assessment data, analysis of secondary data from large national records, and nested data observations
- It generates tables, variables, weights, and other objects to aid large scale assessment data analysis
- It relies on libraries such as: dplyr, stringr, purrr

# Pendings

- add sample data to show functions examples
- add a viggnete
- add merge function

# Done

- add functions for weights in mixed models, methods a and method b, also called normalized and effective sample size weights
- add mean score within observations
- add check cluster id function


## Installation

You can install, in R, the development version of r4sda from this [github](https://github.com/dacarras/r4sda) repository with:

``` r
devtools::install_github("dacarras/r4sda")
```

## List of functions

- `wide_resp()` generates a table of items as rows, and response value as columns, and displays the percentage of responses per item.
- `wide_var()` generates a table of response as rows, and variablesas columns, and displays the percentage of responses per item.
- `stack_resp()` generates a table of items as rows, responses as attributes of items, and the percentage of each response category.
- `c_mean()` estimate cluster means to aid cluster mean centering in mixed models
- `c_wmean()` estimate cluster means to aid cluster mean centering in
    mixed models, including weights within clusters
- `c_sum()` estimate cluster sums to aid cluster variables generation for mixed models
- `reverse()` it generates a reverse score for any given numeric vector (it removes the labels if the variable is labelled).
- `z_score()` it standardize variables returning these as z scores
- `mean_score()` it create mean score of variable (i.e. row wise means)
- `lsa_weights()` add normalized and effective sample weights to the provided data frame
- `senate_weights()` add senate weights scaled up a to a number (e.g. 500, 1000 or else)
- `jkr_iccs()` add jackknifes replicate weights to ICCS 2009 study data frame
- `local_path()` aids the generation of relative working folders, by turning relative locations into a absolute folder paths. Useful for generaing code that works in Unix and Windows machines.
- `wide_resp()` generates a table of items as rows, and response value as columns, and displays the percentage of responses per item.
- `variable_label()` it gets variable labels from a variable, from labelled vector.
- `value_label()` it generate a table from the value labels from a labelled vector.
- `variables_table()` generates a table of a data frame, including variable names, variable types, sample values, and variable labels
- `check_cluster_id()` it tests if the cluster id is unique across the data frame, or if these repeats between addtional cluster factors. For example, it checks if schools id repeats between country observations, or if observations id are unique between schools.
- [...]


## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
# create an items data frame
# items_data <- dplyr::select(data_frame, LS2T01:LS2T16)

```



```text
# request a wide response data table with wide_resp() function
r4sda::wide_resp(items_data)

   variable    `1`    `2`    `3`    `4`    `NA`
   <chr>     <dbl>  <dbl>  <dbl>  <dbl>   <dbl>
 1 LS2T01   0.105  0.308  0.0917 0.486  0.00961
 2 LS2T02   0.697  0.103  0.150  0.0389 0.0104 
 3 LS2T03   0.133  0.0316 0.758  0.0708 0.00664
 4 LS2T04   0.100  0.115  0.0423 0.734  0.00841
 5 LS2T05   0.567  0.0802 0.200  0.147  0.00641
 6 LS2T06   0.276  0.219  0.232  0.250  0.0234 
 7 LS2T07   0.0951 0.213  0.494  0.190  0.00871
 8 LS2T08   0.138  0.237  0.584  0.0317 0.00945
 9 LS2T09   0.191  0.426  0.104  0.270  0.00808
10 LS2T10   0.382  0.187  0.202  0.209  0.0208 
11 LS2T11   0.447  0.271  0.107  0.162  0.0125 
12 LS2T12   0.274  0.242  0.159  0.309  0.0150 
13 LS2T13   0.247  0.417  0.149  0.170  0.0166 
14 LS2T14   0.600  0.159  0.0940 0.129  0.0187 
15 LS2T15   0.0956 0.134  0.0665 0.683  0.0210 
16 LS2T16   0.0801 0.593  0.0948 0.208  0.0240 
```

Because this is a table, it can be render as kable table: 

You can also embed plots, for example:

```text
knitr::kable(wide_resp(items_data), digits = 2)

```

|variable |    1|    2|    3|    4|   NA|
|:--------|----:|----:|----:|----:|----:|
|LS2T01   | 0.11| 0.31| 0.09| 0.49| 0.01|
|LS2T02   | 0.70| 0.10| 0.15| 0.04| 0.01|
|LS2T03   | 0.13| 0.03| 0.76| 0.07| 0.01|
|LS2T04   | 0.10| 0.12| 0.04| 0.73| 0.01|
|LS2T05   | 0.57| 0.08| 0.20| 0.15| 0.01|
|LS2T06   | 0.28| 0.22| 0.23| 0.25| 0.02|
|LS2T07   | 0.10| 0.21| 0.49| 0.19| 0.01|
|LS2T08   | 0.14| 0.24| 0.58| 0.03| 0.01|
|LS2T09   | 0.19| 0.43| 0.10| 0.27| 0.01|
|LS2T10   | 0.38| 0.19| 0.20| 0.21| 0.02|
|LS2T11   | 0.45| 0.27| 0.11| 0.16| 0.01|
|LS2T12   | 0.27| 0.24| 0.16| 0.31| 0.02|
|LS2T13   | 0.25| 0.42| 0.15| 0.17| 0.02|
|LS2T14   | 0.60| 0.16| 0.09| 0.13| 0.02|
|LS2T15   | 0.10| 0.13| 0.07| 0.68| 0.02|
|LS2T16   | 0.08| 0.59| 0.09| 0.21| 0.02|

>Note: this readme is provisory
